An elaborate data set on human gait and the effect of mechanical perturbations

### 2014-12-14 paper.tex
Proofed the perturbation signals section.

\section*{Perturbation Signals}
%
The data set uses a testing protocol containing a segment of normal walking,
followed by a series of pseudo-random longitudinal perturbations, and ending
with a second segment of normal walking. Three pseudo-random speed control
signals, with mean velocities of 0.8, 1.2, and 1.6 $m \cdot s^{-1}$, were
generated through MATLAB and Simulink. Initially, a random acceleration signal
was created using discrete-time Gaussian white noise and saturated at the
maximum belt acceleration of 15 $m \cdot s^{-2}$, integrated to obtain belt
speed, and high-pass filtered (second-order Butterworth) to eliminate velocity
drift. Mean velocity was then added to the signal and limited between 0 and 3.6
m/s. The cutoff frequencies of the high-pass filter, as well as the variance in
the acceleration signal, were adjusted until the appropriate standard
deviations for each mean velocity were obtained.

These standard deviations were determined through a trial and error method.
Initially, perturbations of $\pm$ 10\% for each mean velocity were generated.
However, this large deviation in the slowest mean velocity (0.8 $m \cdot
s^{-1}$) would unintentionally reduce the treadmill speed to values close to
zero. This would produce an undesired stopping and starting sensation.
Alternatively, smaller standard perturbations of $\pm$ 5\% did not produce
noticeable perturbations in the fastest trial (1.6 $m \cdot s^{-1}$).
Therefore, standard deviations of 7.16\%, 10.23\%, and 13.1\% were selected for
the 0.8, 1.2, and 1.6 $m \cdot s^{-1}$ speeds, respectively.  This ensured that
the test subject was sufficiently perturbed for each speed, while remaining
within the limits of our testing protocol.

To ensure that the treadmill could follow the higher frequencies of the random
signals, the maximum treadmill acceleration of 15 $m \cdot s^{-2}$ was used in
the D-Flow Software. As this value remains constant throughout the experiment,
the initial starting and stopping of the treadmill would be too rapid for test
subjects. To eliminate this, a ramped acceleration and deceleration was
generated and placed at the beginning and end of the entire signal,
respectively. Other additions include the one minute segments of unperturbed
walking at the beginning and end of the trial and 30 seconds of no movement for
force plate zeroing and subject T-pose calibration.  The sampling rate of the
simulation was set to 100 Hz, corresponding to the D-Flow software's
capabilities.

The resulting output from the MATLAB script and Simulink produces a comma-
delimited text file of six columns: time stamp, slow, normal, and fast walking
perturbation signals, and slow and fast running signals.  These running signals
were generated in the event the testing protocol expanded from a walking study
to a running study, though they were not used in the data set.  The actual
motion of the treadmill belts are compared to the original input signals as
shown in Figure~\ref{fig:input_output}.   The standard deviations of the
outputs were calculated at 6.07\%, 9.68\%, and 12.68\% for the 0.8, 1.2,
and 1.6 $m \cdot s^{-1}$ speeds, respectively.

\begin{figure*}
  \centering
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt velocity input signals (purple) and recorded output
    velocities (blue) for average belt speeds of 0.8, 1.2, and 1.6 m/s,
    respectively}
  \label{fig:input_output}
\end{figure*}

For further analysis, the input and output signals for each average speed were
transformed into the frequency domain through the Fast Fourier algorithm, and
the results are shown in Figure~\ref{fig:freq_analysis}. It was observed that
for the 1.2 $m \cdot s^{-1}$ average walking speed, the amplitude of the output
is significantly lower than the amplitude of the input signal at lower
frequencies. Additionally, the amplitude of output signal in the 0.8 $m \cdot
s^{-1}$ walking speed begins to attenuate around 2 Hz, which is a noticeably
lower frequency than the other walking speeds.

\begin{figure*}
  \centering
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency spectrum of the treadmill belt velocity input signal
    (purple) and the recorded output velocity (blue) for average belt speeds of
    0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure*}

@@

\begin{listing*}
  \begin{minted}[gobble=4]{python}
    >>> from gaitanalysis.motek import DFlowData
    >>> data = DFlowData('mocap-020.txt', 'record-020.txt', 'meta-020.yml')
    >>> mass = data.meta['subject']['mass']
    >>> data.clean_data()
    >>> event_df = dflow_data.extract_processed_data(event='Longitudinal Perturbation')
  \end{minted}
  \cprotect\caption{Python interpreter session showing how one could load a
    trial into memory, extract the subject's mass from the meta data, run the
    data cleaning process, and finally extract a Pandas \verb|DataFrame|
    containing all of the time histories for a specific event in the trial.}
  \label{lis:dflowdata}
\end{listing*}

>>>>

\section*{Perturbation Signals}
%
As previously described, the protocol included a phase of normal walking,
followed by longitudinal belt speed perturbations, and ended with a second
segment of normal walking. Three pseudo-random belt speed control signals, with
mean velocities of \SIlist{0.8;1.2;1.6}{\meter\per\second}, were pre-generated
with MATLAB and Simulink (Mathworks, Natick, Massachusetts, USA).  The same
control signal was used for all trials at that given speed.

To create the signals, we started by generating random 100~\si{\hertz}
acceleration signals using the Simulink discrete-time Gaussian white noise
block followed by a saturation block set at the maximum belt acceleration of
15~\si{\meter\per\second\squared}. The signal was then integrated to obtain
belt speed and high-pass filtered with a second-order Butterworth filter to
eliminate drift. One of the three mean speeds were then added to the signal and
limited between \SIrange{0}{3.6}{\meter\per\second}. The cutoff frequencies of
the high-pass filter, as well as the variance in the acceleration signal, were
manually adjusted until acceptable standard deviations for each mean speed were
obtained: \SIlist{0.06;0.12;0.21}{\meter\per\second} for the three speeds,
respectively. These ensured that the test subjects were sufficiently perturbed
at each speed, while remaining within the limits of our equipment and testing
protocol.

To ensure that the treadmill belts could accelerate to the desired values, the
high performance mode in the D-Flow software was enabled. This had the side
effect of enabling too rapid of accelerations when the belt speed changed to or
from zero speed. To eliminate this, a suitable ramped acceleration and
deceleration were generated for the speed transitions.

The MATLAB script and Simulink model produce a comma-delimited text file of six
signals: time stamp, slow, normal, and fast walking perturbation signals, and
slow and fast running signals.~\footnote{The running signals were not used in
the experiments presented in this paper.} The measured speed of the treadmill
belts are compared to the control input signals in
Figure~\ref{fig:input_output} to show the effect of the treadmill and
controller dynamics. The system introduces a delay and seems to act as a low
pass filter. The standard deviations of the outputs do not significantly differ
from the desired values: \SIlist{0.05;0.12;0.2}{\meter\per\second} for the
three speeds, respectively.
%
\begin{figure}
  \centering
  \includegraphics{figures/input_vs_output.pdf}
  \caption{Treadmill belt speed input signals (purple) and recorded output
    speeds (blue) for average belt speeds of
    \SIlist{0.8;1.2;1.6}{\meter\per\second}, respectively.}
  \label{fig:input_output}
\end{figure}

To show the effects of the treadmill dynamics and give an idea of the frequency
content of the actual perturbations, the input and output for each speed were
transformed into the frequency domain using the Fast Fourier algorithm, and the
results are shown in Figure~\ref{fig:freq_analysis}. This shows that for the
1.2~\si{\meter\per\second} walking speed, the amplitude of the output is
significantly lower than the amplitude of the input signal at lower
frequencies. Additionally, the amplitude of output signal in the
0.8~\si{\meter\per\second}  walking speed begins to attenuate around
2~\si{\hertz}, which is a noticeably lower frequency than the other walking
speeds.
%
\begin{figure}
  \centering
  \includegraphics{figures/frequency_analysis.pdf}
  \caption{Frequency spectrum of the treadmill belt velocity input signal
    (purple) and the recorded output velocity (blue) for average belt speeds of
    0.8, 1.2, and 1.6 m/s, respectively}
  \label{fig:freq_analysis}
\end{figure}

@@

\begin{listing}
  \begin{minted}[gobble=4]{python}
    >>> from gaitanalysis.motek import DFlowData
    >>> data = DFlowData('mocap-020.txt', 'record-020.txt', 'meta-020.yml')
    >>> mass = data.meta['subject']['mass']
    >>> data.clean_data()
    >>> event_df = dflow_data.extract_processed_data(event='Longitudinal Perturbation')
  \end{minted}
  \cprotect\caption{Python interpreter session showing how one could load a
    trial into memory, extract the subject's mass from the meta data, run the
    data cleaning process, and finally extract a Pandas \verb|DataFrame|
    containing all of the time histories for a specific event in the trial.}
  \label{lis:dflowdata}
\end{listing}
